---
title: "p8105_hw2_js5095"
author: Jianting Shi
date: 2020-09-27
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(tidyr)
```

## Problem 1

Read the Mr. Trashwheel dataset

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read the precipitation data of 2017 and 2018

```{r}
precip_2017 =
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)

precip_2018 =
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)
```

Combine annual precipitation

```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2017,precip_2018)

left_join(precip_df, month_df, by = "month") %>%
  relocate(month_name, .after = "year")
```

This dataset 'r 

## Problem 2

### 2.1 Read the NYC Transit dataset

```{r}
NYCTransit_df = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

The NYC Transit dataset now contains the following variables: `r names(NYCTransit_df)`. \
The data cleaning steps includes extraction of `r ncol(NYCTransit_df)` columns out of the original dataset with names modified to lowercase and space replaced with "_", as well as the conversion of "entry" variables from character to logical. \
Now the resulting dataframe contains `r nrow(NYCTransit_df)` rows and `r ncol(NYCTransit_df)` columns.

### 2.2 Question 1: How many distinct stations are there?

```{r}
distinct_station = (
  NYCTransit_df %>%
 distinct(station_name, line) %>% 
 nrow()
)
```

There are `r distinct_station` distinct stations.

### 2.3 Question 2: How many stations are ADA compliant?

```{r}
ADA_station = (
  filter(NYCTransit_df, ada == TRUE) %>%
  distinct(station_name, line) %>%
  nrow()
)
```

There are `r ADA_station` ADA compliant stations.

### 2.4 Question 3: What proportion of station entrances / exits without vending allow entrance?

```{r}
entry_no_vending_entrance = (
  NYCTransit_df %>%
  filter(entry == TRUE & vending == "NO") %>%
  nrow()
)

no_vending_entrance = (
  NYCTransit_df %>%
  filter(vending == "NO") %>%
  nrow()
)

prop = entry_no_vending_entrance/no_vending_entrance
```

The proportion is `r prop`.

### 2.5 Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? 

```{r}
NYCTransit_df %>% 
  
  pivot_longer(route1:route11, names_to = "route_number", values_to = "route_name") %>%
  filter(route_name == 'A') %>% 
  distinct(station_name, line) %>% 
  nrow()
```


### 2.6 Of the stations that serve the A train, how many are ADA compliant?

```{r}
NYCTransit_df %>% 
  pivot_longer(route1:route7, names_to = "route_number", values_to = "route_name") %>%
  filter(route_name == 'A', ada == TRUE) %>% 
  distinct(station_name, line) %>% 
  nrow()
```


# Problem 3

Read the FiveThirtyEight data

```{r}
polsmonth_df = 
  read_csv(
    "./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year","month", "day")) %>%
  mutate(month = as.numeric(as.factor(month))) %>%
  mutate(month = month.name[month]) %>%
  mutate(president = as.character(prez_gop)) %>%
  mutate(president = recode(president, "0" = "dem", "1" = "gop"))
```


